---
title: "Applied GLM Final Project"
author: "Lauren Bergam"
date: "April 21, 2019"
output: pdf_document
---

```{r}
library(ggplot2)

load("alzheimers_data_filtered.Rdata")
nacc <- alzheimers_data_filtered

```

```{r}
# Clean data, errant value for apoe_01
nacc <- nacc[(nacc$apoe_01 == 0 | nacc$apoe_01 == 1), ] 

# Contingency table for apoe and dementia diagnosis 
table_apoe01 <- table("APOE Carrier" = factor(nacc$apoe_01, levels = c("1", "0")), 
                    "Dementia" = factor(nacc$DEMENTED, levels = c("1", "0"))) 
table_apoe01

# Fisher's test
fisher.test(table_apoe01)

```
Interpretation: The odds of having dementia is 2.3 times more likely for individuals who are a carrier of the apolipoprotein E than those who are non-carriers. The chi-squared test gives a very small p-value and therefore, dementia and apoe carriers are not independent from each other.

```{r}
# Contingency table for age
nacc["age_threshold"] = ifelse(nacc$age_visit > 70, 1, 0)
age_table <- table("Age > 70" = factor(nacc$age_threshold, levels = c("1", "0")), 
                   "Dementia" =  factor(nacc$DEMENTED, levels = c("1", "0")))
age_table

# Fisher's test
fisher.test(age_table)

```
Since age is a discrete, non-binary variable, we choose a split value (in this case 70) to classify the data into two groups for a contingency table. 
Interpretation: The odds of having dementia is 21.6 times more likely for individuals over the age of 70 than those who are equal to or under the age of 70. The chi-squared test gives a very small p-value and therefore, dementia and age are not independent from each other. 

```{r, cache = TRUE}
# Contingency table for animal test
nacc["animal_threshold"] = ifelse(nacc$animal < 10, 1, 0)
animal_table <- table("Animal < 10 " = factor(nacc$animal_threshold, levels = c("1", "0")), 
                      "Dementia" =  factor(nacc$DEMENTED, levels = c("1", "0")))
animal_table

# Fisher's Test 
fisher.test(animal_table)

```
Since number of animals recalled is a discrete, non-binary variable, we choose a split value (in this case 10) to classify the data into two groups. 
Interpretation: The odds of having dementia is 29.8 times more likely for individuals who can recall fewer than 10 animals within one minute than those who can recall 10 or more. The chi-squared test gives a very small p-value and therefore, dementia and the results from the animal test are not independent from each other. 

```{r, cache = TRUE}
# Contingency table for years of education
nacc["educ_threshold"] = ifelse(nacc$educ < 12, 1, 0)
educ_threshold <- table("# of years of education < 12 " = factor(nacc$educ_threshold, levels = c("1", "0")), 
                        "Dementia" =  factor(nacc$DEMENTED, levels = c("1", "0")))
educ_threshold


# Fisher's Test
fisher.test(educ_threshold)

# Chi-squared test
chisq.test(educ_threshold)

```
Interpretation: The odds of having dementia is 2.2 times more likely for individuals who have fewer than 12 years of education than those who have 8 years or more. The chi-squared test gives a very small p-value and therefore, dementia and years of education are not independent from each other.

```{r, cache = TRUE}
# 6x2 contingency table for APOE genotype
apoe_table <- table(apoe_genotype = factor(nacc$apoe), Dementia = factor (nacc$DEMENTED))
apoe_table

# Convert to dataframe with frequencies for each permutation
apoe_df <- as.data.frame(apoe_table)
apoe_df

# Plot frequency table for each genotype
ggplot(apoe_df, aes(x = apoe_genotype, y = Freq, fill = Dementia)) + 
  geom_bar(position = "dodge", stat = "identity") +
  geom_text(aes(label = Freq), vjust = -0.5, size = 3, position = position_dodge(0.9))

# Model the log odds of having dementia for each genotype compared to the baseline of genotype 1
apoe_model <- glm(Dementia ~ apoe_genotype, weights = Freq, data = apoe_df, family = binomial)
summary(apoe_model)

exp(-3.77182)*100
exp(-3.77182+0.96613)*100
exp(-3.77182-0.05682)*100

```
Interpretation: The intercept is the log odds of having dementia given the patient has genotype 1. Exponentiating shows that the odds of having dementia given genotype 1 are 2.3%. The odds of the other genotypes are calculated relative to this baseline. The highest odds of having dementia given any of the genotypes is 6.0% for genotype 4 and the lowest odds is 2.1% for genotype 6. 








