---
title: "AGLM Project Presentation"
author: "Peter Shewmaker"
date: "April 21, 2019"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(lmtest)
load(file = "alzheimers_data_filtered.Rdata")
```

```{r, echo = FALSE}
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```

## Nested Models

- Two models are called \textit{nested} if one is a special case of the other.
- For example, if:

$$Y_1 = \alpha_1 + \beta_2X_2$$
$$Y_2 = \alpha_1 + \beta_1X_1 + \beta_2X_2 + \beta_3X_3$$
then the first model is nested within the second model. This is because $Y_1 = Y_2$ if both $X_1$ and $X_3$ are held at 0.

- If two models are nested, we can perform a Likelihood Ratio Test to determine which of the models may be a better fit.

## Likelihood Ratio Test

- Null hypothesis: the simpler model is the better one.

- The test statistic is a constant multiple of the logarithm of the likelihood ratio of the simpler (smaller) model to the log-likelihood of the larger model, that is:
$$ LRT = -2\log\left(\frac{\mathcal{L}_s(\theta)}{\mathcal{L}_g(\theta)}\right) = -2[\log(\mathcal{L}_s(\theta) - \log(\mathcal{L}_g(\theta))]$$
where $\mathcal{L}_s(\theta)$ is the likelihood of the smaller model, and $\mathcal{L}_g(\theta)$ is the likelihood of the larger model.

- The test statistic approximates a chi-squared random variable, with degrees of freedom equal to the difference of the number of parameters between the two models.

- The function "lrtest" in the package "lmtest" performs the likelihood ratio test for two nested (generalized) linear models.

## Predicting memory score with a linear model

- The variable "log_mem_imm" is a measure of the patient's memory. The following models use some demographic information as well as performance on cognitive tests to predict the patient's score on this memory test.
```{r, echo = TRUE}
m1 <- lm(log_mem_imm ~ age_visit + mmse + cdr_global, 
         data = alzheimers_data_filtered)
m2 <- lm(log_mem_imm ~ age_visit + mmse + cdr_global
         + sex + ethnicity + educ,
         data = alzheimers_data_filtered)
```

## Performing a Likelihood Ratio Test with "lrtest"
 
```{r, echo = TRUE, size = "tiny"}
lrtest(m1, m2)
```

- Since the test statistic is sufficiently large, we can reject the null hypothesis and conclude that the larger model is more likely than the smaller one in this case.

